# Demographics

## Installation
```{r installation_demog}
library(acsprojectns)
library(devtools)
load_all()
library(icdb)
library(kableExtra)
```
## Index ACS event
Get the index ACS event
```{r demog_index_acs, results="markup"}
# The index event
index_acs <- msrv$sus$apc_spells_diagnosis |>
  # Time window
  dplyr::filter(spell_start >= acs_date_min & spell_end <= acs_date_max) |>
  # # Emergency admission
  dplyr::filter(as.character(method_of_admission_hospital_provider_spell) %in% !!adm_method$admission_method_code) %>% 
  # Only cohort NHS numbers
  {if(.Platform$OS.type == "windows") {
    dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
  } else {
    dplyr::filter(., nhs_number %in% cohort_ids)
  }} |>
  # Collect to local
  icdb::run() |>
  # Recode the ICD10 codes
  dplyr::mutate(nhs_number = as.character(nhs_number), 
                primary_diagnosis = icdb::icd10(primary_diagnosis_icd, 
                                                system.file("icd10_codes", "icd10_ns.yaml", package="acsprojectns"))) %>% 
  # Make sure ICD10 codes are all valid
  # {if(all(icdb::is_valid(.$primary_diagnosis))) . else stop("error1 - couldn't parse all the ICD codes")} |>
  # Only ACS events TODO: change this to filter
  dplyr::filter(primary_diagnosis %in_group% c("acs_stemi", "acs_nstemi", "acs_unstable_angina")) |>
  # Only the first ACS event
  dplyr::group_by(nhs_number) |>
  dplyr::slice_min(spell_start, with_ties=F) |>
  # always ungroup stuff
  dplyr::ungroup()

# Tests
stopifnot(nrow(index_acs)==length(cohort_ids),
          all(cohort_ids %in% index_acs$nhs_number))

# View
kbl(index_acs |> dplyr::slice_head(n=5), longtable = T, booktabs = T, caption = "Index ACS event")
```

## Age
Get the age of the patient at the index ACS event
```{r demographics_age, results="markup"}
dat_age <- index_acs |> dplyr::select(nhs_number, age = age_on_admission)
```

## Sex
```{r demographics_sex}
# read in the code mappnigs for sex
sex_codes <- read.csv(system.file("sex_codes/sex_codes.csv", package = "acsprojectns"))

# get the sex data from the SUS data, then add in that from the SWD attribute history data
dat_sex <- msrv$sus$apc_spells_primary_diagnosis %>% 
  # Only cohort NHS numbers
  {if(.Platform$OS.type == "windows") {
    dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
  } else {
    dplyr::filter(., nhs_number %in% cohort_ids)
  }} |>
  # only nhs number and sex
  dplyr::select(nhs_number, sex_sus = sex) |>
  # combine with the sex data from the SWD attribute history
  dplyr::left_join(msrv$swd$attr_h %>% 
                       # Only cohort NHS numbers
                      {if(.Platform$OS.type == "windows") {
                        dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
                      } else {
                        dplyr::filter(., nhs_number %in% cohort_ids)
                      }} |>
                     dplyr::select(nhs_number, sex_swd = sex), by="nhs_number") |>
  # collect to local
  run() |>
  # ensure all characters
  dplyr::mutate(dplyr::across(dplyr::starts_with("sex"), ~as.character(.x))) |>
  # combine into one column
  tidyr::pivot_longer(dplyr::starts_with("sex"), names_to="source", values_to="sex_code") |>
  # recode according to the mappings
  dplyr::left_join(sex_codes, by="sex_code") |>
  # group by each patient 
  dplyr::group_by(nhs_number) |>
  # take the mode within each patient, ignoring NAs
  dplyr::summarise(sex = acsprojectns::mode(sex)) |>
  # shouldnt need to do this at work 
  dplyr::mutate(nhs_number = as.character(nhs_number)) |>
  # ensure all nhs numbers represented
  dplyr::right_join(index_acs |> dplyr::select(nhs_number), by="nhs_number")
```

## Ethnicity
```{r demographics_ethnicity, results="markup"}
# read in the ethnicity mappings
ethnicity_codes <- read.csv(system.file("ethnicity_codes/ethnicity_codes.csv", package = "acsprojectns"))

# get the ethnicity data from the SUS data, then add in that from the SWD attribute history data
dat_ethnicity <- msrv$sus$apc_spells_primary_diagnosis %>% 
  # Only cohort NHS numbers
  {if(.Platform$OS.type == "windows") {
    dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
  } else {
    dplyr::filter(., nhs_number %in% cohort_ids)
  }} |>
  # Just ethnicity
  dplyr::select(nhs_number, ethnicity_sus = ethnic_group) |>
  # Add SWD attribute data
  dplyr::left_join(msrv$swd$attr_h  %>% 
                      # Only cohort NHS numbers
                      {if(.Platform$OS.type == "windows") {
                        dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
                      } else {
                        dplyr::filter(., nhs_number %in% cohort_ids)
                      }} |>
                     dplyr::select(nhs_number, ethnicity_swd = ethnicity), by="nhs_number") |>
  # collect to local
  icdb::run() |>
  # ensure all characters
  dplyr::mutate(dplyr::across(dplyr::starts_with("ethnicity"), ~as.character(.x))) |>
  # combine into one column
  tidyr::pivot_longer(dplyr::starts_with("ethnicity"), names_to="source", values_to="ethnicity_code") |>
  # recode according to the mappings
  dplyr::left_join(ethnicity_codes, by="ethnicity_code") |>
  # group by each patient 
  dplyr::group_by(nhs_number) |>
  # take the mode within each patient, ignoring NAs; n.b. use 'ethnicity' rather than
  # 'ethnicity_group' for more detailed breakdown
  dplyr::summarise(ethnicity = acsprojectns::mode(ethnicity_group)) |>
  # shouldnt need to do this at work 
  dplyr::mutate(nhs_number = as.character(nhs_number)) |>
  # ensure all nhs numbers represented
  dplyr::right_join(index_acs |> dplyr::select(nhs_number), by="nhs_number")
  
# View 
kbl(dplyr::slice_head(dat_ethnicity, n=5), longtable = T, booktabs = T, caption = "Ethnicity data")
```

## Index of Multiple Deprivation
```{r demographics_imd, results="markup"}
# extract the IMD data from the swd_attributes table
dat_imd <- msrv$swd$swd_attribute %>% 
  # Only cohort NHS numbers
  {if(.Platform$OS.type == "windows") {
    dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
  } else {
    dplyr::filter(., nhs_number %in% cohort_ids)
  }} |>
  # just the IMD
  dplyr::select(nhs_number, wd_imd_decile_19) |>
  # collect to local 
  icdb::run()  |>
  # group by each patient 
  dplyr::group_by(nhs_number) |>
  # take the most recent
  dplyr::slice_head(n=1) |>
  # shouldnt need to do this at work 
  dplyr::mutate(nhs_number = as.character(nhs_number)) |>
  # ensure all nhs numbers represented
  dplyr::right_join(index_acs |> dplyr::select(nhs_number), by="nhs_number")

# View 
kbl(dplyr::slice_head(dat_imd, n=5), longtable = T, booktabs = T, caption = "IMD data")
```

## Employment decile
```{r demographics_employment, results="markup"}
# extract the employment decile data from the swd_attributes table
dat_employment_decile <- msrv$swd$swd_attribute %>% 
  # Only cohort NHS numbers
  {if(.Platform$OS.type == "windows") {
    dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
  } else {
    dplyr::filter(., nhs_number %in% cohort_ids)
  }} |>
  # just the employment decile
  dplyr::select(nhs_number, wd_employment_decile) |>
  # collect to local 
  icdb::run() |>
  # group by each patient 
  dplyr::group_by(nhs_number) |>
  # take the most recent
  dplyr::slice_head(n=1) |> 
  # shouldnt need to do this at work 
  dplyr::mutate(nhs_number = as.character(nhs_number)) |>
  # ensure all nhs numbers represented
  dplyr::right_join(index_acs |> dplyr::select(nhs_number), by="nhs_number")

# View 
kbl(dplyr::slice_head(dat_employment_decile, n=5), longtable = T, booktabs = T, caption = "Employment decile data")
```

## Income decile
```{r demographics_income, results="markup"}
# extract the income decile data from the swd_attributes table
dat_income_decile <- msrv$swd$swd_attribute %>% 
  # Only cohort NHS numbers
  {if(.Platform$OS.type == "windows") {
    dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
  } else {
    dplyr::filter(., nhs_number %in% cohort_ids)
  }} |>
  # just the income decile
  dplyr::select(nhs_number, wd_income_decile) |>
  # collect to local 
  icdb::run() |>
  # group by each patient 
  dplyr::group_by(nhs_number) |>
  # take the most recent
  dplyr::slice_head(n=1) |>
  # shouldnt need to do this at work 
  dplyr::mutate(nhs_number = as.character(nhs_number)) |>
  # ensure all nhs numbers represented
  dplyr::right_join(index_acs |> dplyr::select(nhs_number), by="nhs_number")

# View 
kbl(dplyr::slice_head(dat_income_decile, n=5), longtable = T, booktabs = T, caption = "Income decile data")
```

## Frailty
```{r demographics_frailty, results="markup"}
# extract the electronic frailty index data from the swd attributes history table
dat_frailty <- msrv$swd$attr_h %>% 
  # Only cohort NHS numbers
  {if(.Platform$OS.type == "windows") {
    dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
  } else {
    dplyr::filter(., nhs_number %in% cohort_ids)
  }} |>
  # just the efi decile
  dplyr::select(nhs_number, attribute_period, score = efi_category) |>
  # label the score
  dplyr::mutate(frailty_score = "efi") |>
  # collect to local 
  icdb::run() |>
  # add the Cambridge segment data
  dplyr::bind_rows(msrv$swd$cambridge %>% 
                       # Only cohort NHS numbers
                      {if(.Platform$OS.type == "windows") {
                        dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
                      } else {
                        dplyr::filter(., nhs_number %in% cohort_ids)
                      }} |>
                     dplyr::select(nhs_number, attribute_period, score = segment) |>
                     dplyr::mutate(frailty_score = "cambridge_segment") |> 
                     icdb::run()) |>
  # add the Cambridge score data
  dplyr::bind_rows(msrv$swd$cambridge %>% 
                      # Only cohort NHS numbers
                      {if(.Platform$OS.type == "windows") {
                        dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
                      } else {
                        dplyr::filter(., nhs_number %in% cohort_ids)
                      }} |>
                     dplyr::select(nhs_number, attribute_period, score = cambridge_score) |>
                     dplyr::mutate(frailty_score = "cambridge_score") |> 
                     icdb::run()) |>
  # shouldnt need to do this at work 
  dplyr::mutate(nhs_number = as.character(nhs_number)) |>
  # join with the index ACS event date, so that we can filter for data existing prior to the ACS event
  dplyr::left_join(index_acs |> dplyr::select(nhs_number, spell_start), by="nhs_number") |>
  # only look at data prior to the ACS event
  dplyr::filter(attribute_period < spell_start) |>
  # NAs aren't useful
  dplyr::filter(!is.na(score)) |>
  # take the most recent of each score
  dplyr::group_by(nhs_number, frailty_score) |>
  dplyr::slice_max(attribute_period, with_ties=FALSE) |>
  # select only those we need
  dplyr::select(nhs_number, frailty_score, score) |>
  # separate efi and Cambridge
  tidyr::pivot_wider(names_from=frailty_score, values_from=score) |>
  # shouldnt need to do this at work 
  dplyr::mutate(nhs_number = as.character(nhs_number)) |>
  # ensure all nhs numbers represented
  dplyr::right_join(index_acs |> dplyr::select(nhs_number), by="nhs_number")


# due to ACS events before the start of data collection regarding the Cambridge
# score, there is 40% missingness. Therefore have a look at what the recorded
# score is at either: just before the ACS, or if missing, the monthly extract in
# closest proximity to the ACS (this will be after their ACS event, but should
# give us a good idea of the cormorbidity)
dat_fudge_cambridge <- msrv$swd$cambridge %>% 
  # Only cohort NHS numbers
  {if(.Platform$OS.type == "windows") {
    dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
  } else {
    dplyr::filter(., nhs_number %in% cohort_ids)
  }} |>
  dplyr::select(nhs_number, attribute_period, cambridge_score_coal=cambridge_score, 
                                              cambridge_segment_coal=segment) |>
  icdb::run() |>
  # shouldnt need to do this at work 
  dplyr::mutate(nhs_number = as.character(nhs_number)) |>
  # take the oldest recent of each score
  dplyr::group_by(nhs_number) |>
  dplyr::slice_min(attribute_period, with_ties=FALSE) |>
  # add in the scores from just before the ACS event (from above)
  dplyr::right_join(dat_frailty |> dplyr::select(nhs_number, cambridge_score, cambridge_segment), by="nhs_number") |>
  dplyr::mutate(fudge_cambridge_score   = dplyr::coalesce(cambridge_score, cambridge_score_coal), 
                fudge_cambridge_segment = dplyr::coalesce(cambridge_segment, cambridge_segment_coal)) |>
  # select only those we need
  dplyr::select(nhs_number, fudge_cambridge_score, fudge_cambridge_segment)


# View 
kbl(dplyr::slice_head(dat_frailty, n=5), longtable = T, booktabs = T, caption = "Frailty score data")
```

## Comorbidities
```{r demographics_comorbidities, results="markup"}
# extract all of the hospital admission for the cohort (use episodes for finer details - not sure if
# this makes a difference)
dat_comorbs <- msrv$sus$apc_episodes %>% 
  # Only cohort NHS numbers
  {if(.Platform$OS.type == "windows") {
    dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
  } else {
    dplyr::filter(., nhs_number %in% cohort_ids)
  }} |>
  # collect to local
  icdb::run() |>
  # ensure all columns characters
  dplyr::mutate(dplyr::across(dplyr::matches("primary_diagnosis_icd|secondary_diagnosis_[0-9]+_icd"),
                              ~ as.character(.x))) |>
  # pivot diagnoses to long format
  tidyr::pivot_longer(dplyr::matches("primary_diagnosis_icd|secondary_diagnosis_[0-9]+_icd"),
                      names_to="diag_pos",
                      values_to="diagnosis", 
                      values_drop_na=TRUE) |>
  # don't need to know where the diagnosis appeared in the list, deselect/clean up
  dplyr::select(-diag_pos) |>
  # make sure dates are dates
  dplyr::mutate(dplyr::across(c(episode_start, episode_end), ~as.Date(.x))) |>
  # add the cormorbidities from the SWD
  dplyr::bind_rows(
    msrv$swd$attr_h %>% 
       # Only cohort NHS numbers
      {if(.Platform$OS.type == "windows") {
        dplyr::right_join(., data.frame("nhs_number" = cohort_ids), by="nhs_number", copy = TRUE) 
      } else {
        dplyr::filter(., nhs_number %in% cohort_ids)
      }} |>
      dplyr::select(nhs_number, 
                    episode_start=attribute_period, 
                    episode_end=attribute_period,
                    dplyr::matches(c("smoking", "^obesity", "^ht$", "^ckd$", "anaemia",
                                      "^stroke$", "^liver", "^af$", "arrhythmia", "ihd", "^hf$", 
                                      "vasc", "diabetes"))) |>
     icdb::run() |>
     # pivot diagnosis flags to long format
     tidyr::pivot_longer(!c(nhs_number, episode_start, episode_end),
                         names_to="diagnosis",
                         values_to="flag",
                         values_transform = list(flag=as.logical),
                         values_drop_na=TRUE) |>
     # filter out the '0' flags as these mean that the patient doesn't have that diagnosis
     dplyr::filter(flag) |> dplyr::select(-flag) |>
     # recode the SWD diagnoses as appropriate ICD-10 codes
     dplyr::mutate(diagnosis = dplyr::case_when(
       diagnosis=="ckd"              ~ "N190", # N19,Unspecified kidney failure
       diagnosis=="liver_alcohol"    ~ "K704", # K704,Alcoholic hepatic failure
       diagnosis=="liver_nafl"       ~ "K760", # K760 'Fatty (change of) liver, not elsewhere classified'
       diagnosis=="liver_other"      ~ "K729", # K729,"Hepatic failure, unspecified"
       diagnosis=="hf"               ~ "I500", # I50,All heart failure,Heart failure
       diagnosis=="ht"               ~ "I100", # I10,Essential Hypertension
       diagnosis=="diabetes_1"       ~ "E100", # E10,All type 1 diabetes mellitus
       diagnosis=="diabetes_2"       ~ "E110", # E11,All type 2 diabetes mellitus
       diagnosis=="pre_diabetes"     ~ "R739", # ICD-10 code for prediabetes is R73. 09.
       diagnosis=="obesity"          ~ "E660", # E660,Obesity due to excess calories
       diagnosis=="anaemia_iron"     ~ "D509", # D509,"Iron deficiency anaemia, unspecified
       diagnosis=="anaemia_other"    ~ "D649", # D649,"Anaemia, unspecified"
       diagnosis=="vasc_dis"         ~ "I739", # I739,PVD unspecified
       diagnosis=="liver_nafl"       ~ "K760", # K760 'Fatty (change of) liver, not elsewhere classified'
       diagnosis=="smoking"          ~ "F171", 
       diagnosis=="af"               ~ "I480", # I48,All AF and flutter
       diagnosis=="arrhythmia_other" ~ "I499", # I499,"Cardiac arrhythmia, unspecified
       diagnosis=="ihd_mi"           ~ "I249", # I249,Unspecified acute ischaemic heart disease
       diagnosis=="ihd_nonmi"        ~ "I259", # I259,"Chronic ischaemic heart disease, unspecified
       diagnosis=="stroke"           ~ "I639") # I639,"Cerebral infarction, unspecified"
     )
   ) |>
  # keep the daignosis string, then parse the ICD10 codes
  dplyr::mutate(diagnosis = icdb::icd10(diagnosis, system.file("icd10_codes", "icd10_ns.yaml", package="acsprojectns")), 
                diagnosis_str = icdb::group_string(diagnosis)) %>% 
  # separate the groups if there are multiple
  tidyr::separate_longer_delim(diagnosis_str, delim=",") |>
  # Make sure ICD10 codes are all valid
  # {if(all(icdb::is_valid(.$diagnosis))) . else stop("error2 - couldn't parse all the ICD codes")}
  # extract the group string 
  # Filter out diagnoses that aren't defined by my groups (defined in icd10.yaml, created with JS'
  # `map-editor` application available on the icdb github under the releases section)
  dplyr::filter(icdb::group_string(diagnosis)!="") |>
  # shouldnt need to do this at work 
  dplyr::mutate(nhs_number = as.character(nhs_number)) |>
  # Add the index ACS dates
  dplyr::left_join(index_acs |> dplyr::select(nhs_number, spell_start, spell_end), by="nhs_number") |>
  # Filter out events after the index event spell start, as these cannot be confidently listed as
  # comorbidities that existed prior to the index ACS event.
  dplyr::filter(episode_end < spell_start) |>
  # group by the patient and the diagnosis groupings
  dplyr::group_by(nhs_number, diagnosis_str) |>
  # take one one of each
  dplyr::slice_head(n=1) |>
  # always ungoup
  dplyr::ungroup() |>
  # to wide format 
  dplyr::mutate(value=TRUE) |>
  tidyr::pivot_wider(id_cols=nhs_number, 
                     names_from="diagnosis_str", 
                     values_from=value, 
                     values_fill=FALSE) |>
  # shouldnt need to do this at work 
  dplyr::mutate(nhs_number = as.character(nhs_number)) |>
  # ensure all nhs_numbers are presented
  dplyr::right_join(index_acs |> dplyr::select(nhs_number), by="nhs_number") |>
  # replace any injected NAs with FALSE
  dplyr::mutate(dplyr::across(dplyr::everything(), ~tidyr::replace_na(.x, FALSE)))

```

## Combined demographics
```{r demographics_combined, results="markup"}
dat_lst <- list(dat_age, dat_sex, dat_ethnicity, dat_employment_decile, dat_income_decile, dat_imd,
                dat_frailty, dat_fudge_cambridge, dat_comorbs)
dat_demographics <- purrr::reduce(dat_lst, dplyr::left_join, by='nhs_number')

kbl(
dat_demographics |>
  dplyr::ungroup() |>
  dplyr::summarise(male = sum(sex=="male", na.rm=T)/dplyr::n(),
                   dplyr::across(dplyr::where(is.numeric), list(missing     = ~sum(is.na(.x)),
                                                                missing_pct = ~sum(is.na(.x))/dplyr::n(),
                                                                mean        = ~mean(.x, na.rm=T),
                                                                sd          = ~sd(.x, na.rm=T))),
                   dplyr::across(dplyr::where(is.logical), list(n           = ~sum(.x, na.rm=T),
                                                                pct         = ~sum(.x, na.rm=T)/sum(!is.na(.x))))) |>
  dplyr::mutate(dplyr::across(dplyr::where(is.numeric), ~round(.x,3))) |>
  t(),
longtable = T, booktabs = T, caption = "Demographics summary data")
```

## Demographics by frailty
The fudged split uses the best available Cambridge score/segment. This will be from data calculated `after` the ACS event in ~40% people. 
```{r demographics_by_frailty, results="markup"}
kbl(
dat_demographics |>
  dplyr::group_by(cambridge_segment) |>
  dplyr::summarise(n_in_segment = dplyr::n(),
                   age = mean(age, na.rm=T),
                   male= sum(sex=="male", na.rm=T),
                   dplyr::across(dplyr::where(is.logical), ~sum(.x)/dplyr::n())) |>
  dplyr::mutate(dplyr::across(dplyr::where(is.numeric), ~round(.x,3))) |>
  t(),
longtable = T, booktabs = T, caption = "Demographics by Cambridge segment summary data")

kbl(
dat_demographics |>
  dplyr::group_by(fudge_cambridge_segment) |>
  dplyr::summarise(n_in_segment = dplyr::n(),
                   age = mean(age, na.rm=T),
                   male= sum(sex=="male", na.rm=T),
                   dplyr::across(dplyr::where(is.logical), ~sum(.x)/dplyr::n())) |>
  dplyr::mutate(dplyr::across(dplyr::where(is.numeric), ~round(.x,3))) |>
  t(),
longtable = T, booktabs = T, caption = "Demographics by (fudged) Cambridge segment summary data")

# Plan 
# demographics at index ACS
#     age, sex, ethnicity, 
#     split by frailty, age, type of ACS
```


